---
title: "open-data-movilidad"
output: 
  html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE)
library(tidyverse)
library(sf)
library(tmap)
library(kableExtra)
tmap_mode("view")
sf::sf_use_s2(FALSE)
```


This repository contains code to access OD data (based on mobile phones locations) from the Ministry of Transport, Mobility and Urban Agenda (Spain). 

The raw data and some information on the method used to collect it can be found in the link below. (Do not use Chromium to open the link or you won't be able to see the data!).

https://www.mitma.gob.es/ministerio/covid-19/evolucion-movilidad-big-data/opendata-movilidad

If this data is periodically updated, it may be worth it to create a data r package.

```{r}
# Read the data
dist_202102 = read.table(gzfile("datos/202102_maestra1_mitma_distrito/20210201_maestra_1_mitma_distrito.txt.gz"), header = TRUE, sep = "|")
head(dist_202102)
# dist_202102_2 = read.table(gzfile("datos/202102_maestra2_mitma_distrito/20210201_maestra_2_mitma_distrito.txt.gz"), header = TRUE, sep = "|")
# names(dist_202102_2)
# muni_202102 = read.table(gzfile("datos/202102_maestra1_mitma_municipio/20210201_maestra_1_mitma_municipio.txt.gz"), header = TRUE, sep = "|")
# names(muni_202102)
# muni_202102_2 = read.table(gzfile("datos/202102_maestra2_mitma_municipio/20210201_maestra_2_mitma_municipio.txt.gz"), header = TRUE, sep = "|")
# names(muni_202102_2)

dist = read.table("relaciones_distrito_mitma.csv", header = TRUE, sep = "|")
# muni = read.table("relaciones_municipio_mitma.csv", header = TRUE, sep = "|")

zonificacion_dist <- sf::st_read(
  "zonificacion_distritos/zonificacion-distritos/distritos_mitma.shp")
# zonificacion_muni <- sf::st_read(
#   "zonificacion_municipios/zonificacion-municipios/municipios_mitma.shp")

cod_pro = read.table("cod_ccaa_pro.csv", header = TRUE, sep = ",")
```

```{r}
# create province and district origin
dist_202102 = dist_202102 %>% 
  # slice_head(n = 10) %>%
  separate(origen, into = c("prov_O", "dist_O"), sep = 2, remove = FALSE)  

# filter potetial cyclable trips (< 10km) and group by district
dist_202102 = dist_202102 %>% 
  filter(distancia == "002-005" | distancia == "005-010") %>% 
  group_by(origen, prov_O, dist_O) %>% 
  summarise("num. trips <10 km"= sum(viajes))

# add geometry and province and region tags
dist_202102 = dist_202102 %>% 
  left_join(zonificacion_dist, by = c("origen" = "ID"))
dist_202102$prov_O = as.numeric(dist_202102$prov_O)
dist_202102 = dist_202102 %>% 
  left_join(cod_pro, by = c("prov_O" = "Cod_Provincia"))

table(dist_202102$Nom_CCAA)

# make map plottable
dist_202102 = st_sf(dist_202102)
```

```{r}
# filter Valencian Country
dist_202102_CV = dist_202102 %>% 
   filter(Nom_CCAA == "COMUNIDAD VALENCIANA")

# plot map
qtm(dist_202102_CV, "num. trips <10 km")
```

*Fig 1: Number of trips under 10 km by district (during February 2021) Valencian Country*

<!-- Cycling growth in Valencia city: a quantitative analysis after seven years of investment in cycling infrastructure (and other sustainable transport measures). -->

<!-- Alternative title: 'From Grand Prix (2013) to European Green Capital (2024)' -->

# Background

Valencia has gone in less than 10 years from being one of the most car-oriented Spanish cities (it hosted the Grand Prix in 2013!) to becoming a benchmark in cycling in the Mediterranean area. 
This rapid change is mainly due to a set of courageous social and environmental transformation policies (e.g. investment in cycling infrastructure, bikesharing system, traffic calming measures, etc.) carried out by the municipal government that came to power in 2015. 
For this and other environmental achievements it has just been nominated [European Green Capital 2024](https://environment.ec.europa.eu/news/valencia-elsinore-and-velenje-win-2024-european-green-city-awards-2022-10-28_en).

Main policy implementations:

- [Cycling Ring](https://energy-cities.eu/best-practice/a-new-cycling-ring-to-transform-mobility-in-valencia/)

- Bikesharing: [Valenbici](https://www.valenbisi.es/en/home)

- [Traffic Calming Measures](https://elpais.com/espana/comunidad-valenciana/2022-08-07/valencia-revoluciona-su-centro-historico-con-la-peatonalizacion-de-30000-metros-cuadros.html)

# Aims

- Estimate cycling growth. To what extent this growth can be associated to the new policies?

- Estimate economic and health benefits. What are the economic and health benefits of this change? How many deaths (emissions and road safety) have been avoided? 

- Future. Is this growth enough? What to do to go beyond?

# Data 

- [Cycling traffic counts](https://opendata.vlci.valencia.es/en/dataset/intensitat-dels-punts-de-mesura-de-bicicletes-espires-electromagnetiques)


```{r}
library("rjson")

json_file <- "https://geoportal.valencia.es/apps/OpenData/Trafico/tra_espiras_bici_p.json"
json_data <- fromJSON(paste(readLines(json_file), collapse=""))
```

- Valenbici data?

# Next steps

- Get more data!